---
title: "Course Reserves Statistics"
author: "Kevin A Thomas"
date: today
execute: 
  echo: false
  warning: false
  error: false
format: pdf
---

```{r}
# Dependencies
library(dplyr)
library(kableExtra)

# Penn colors
colors_penn <- list(primary = c('#011f5b', '#990000'),
                    secondary = c('#3479be', '#faa755', '#bbd86b', '#bd60a5'),
                    gray = c('#555d6a', '#757f8f', '#a7aeba', '#e1e4ec'))
```

```{r}
# Load data
#dat <- read.csv('~/GitHub/course-reserves-stats/res/leganto-usage-views-2026-03-10.csv')
dat <- read.csv('~/GitHub/course-reserves-stats/res/leganto-usage-views-2026-03-17.csv')
#Replace hard-coded filepath with a dynamic selector in a GUI
dat$Citation <- as.character(dat$Citation)
```

```{r}
#| eval: false
# Filter data (applies to all content) in file with all data
course_codes <- c('BAN_PUBH-5010-001 202610', 'BAN_PUBH-5010-001 202520', 'BAN_PUBH-5010-001 202510', 'BAN_PUBH-5010-001 202430', 'BAN_PUBH-5010-001 202420', 'BAN_PUBH-5010-001 202410')
dat <- filter(dat, Course.code %in% course_codes & User.role == 'Student')
```

```{r}
# Filter data (applies to all content) in pre-filtered file
dat <- filter(dat, User.role == 'Student')
```

These data cover courses from `{r} dat$Course.term[dat$Event.date == min(dat$Event.date)][1]` `{r} dat$Course.year[dat$Event.date == min(dat$Event.date)][1]` through `{r} dat$Course.term[dat$Event.date == max(dat$Event.date)][1]` `{r} dat$Course.year[dat$Event.date == max(dat$Event.date)][1]`.

Note: Data do not include readings that were not viewed.

## Summary by viewed item

```{r}
#User-selectable filters (apply to Summary by item section content only): Course year, Course term for the by-item section only
```

To do:

- Fix if chart is longer than the page--'scale_down' seems to ignore the bottom margin
- Sorting by rate seems not to work
- Fix histogram label spacing

```{r}
# Total student views by item (table)
dat_temp <- dat |>
  group_by(Citation) |>
  summarize(Courses = length(unique(Course.code)),
            'Minimum\nenrollment' = min(Course.enrollment),
            'Maximum\nenrollment' = max(Course.enrollment),
            Views = sum(Total.views)
            ) |>
  arrange(Citation)
kbl(dat_temp,
    format = 'latex',
    caption = 'Number of student views by viewed item',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE
    ) |>
  kable_styling(c('striped', 'hover'),
                latex_options = c('HOLD_position', 'scale_down')
                )
```

```{r}
# Most- and least-viewed items (table)
dat_temp <- dat |>
  group_by(Citation, Course.code) |>
  summarize(Enrollment = median(Course.enrollment),
            Views = sum(Total.views)
            ) |>
  group_by(Citation) |>
  summarize(Courses = length(unique(Course.code)),
            'Median\nenrollment' = median(Enrollment),
            Views = sum(Views),
            'Views per\nstudent' = round(sum(Views) / sum(Enrollment), 2)
            ) |>
  arrange(desc(`Views per\nstudent`), .by_group = FALSE)
if(nrow(dat_temp) > 10){
  dat_temp <- rbind(dat_temp[1:5, ],
                    rep("...", ncol(dat_temp)),
                    dat_temp[(nrow(dat_temp)-4):nrow(dat_temp), ]
                    )
}
kbl(dat_temp,
    format = 'latex',
    caption = 'Most-viewed and least-viewed items with views',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE
    ) |>
  kable_styling(c('striped', 'hover'),
                latex_options = c('HOLD_position', 'scale_down')
                )
```
```{r}
# Seldom-viewed items (table)
dat_temp <- dat |>
  group_by(Citation, Course.code) |>
  summarize(Enrollment = median(Course.enrollment),
            Views = sum(Total.views)
            ) |>
  group_by(Citation) |>
  summarize(Courses = length(unique(Course.code)),
            'Median\nenrollment' = median(Enrollment),
            Views = sum(Views),
            'Views per\nstudent' = round(sum(Views) / sum(Enrollment), 2)
            ) |>
  filter(`Views per\nstudent` < 1) |>
  arrange(desc(`Views per\nstudent`), .by_group = FALSE)
if(nrow(dat_temp) == 0){
  dat_temp <- rbind(dat_temp,
                    rep("N/A", ncol(dat_temp))
                    )
}
kbl(dat_temp,
    format = 'latex',
    caption = 'Viewed items with under one view per student',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE
    ) |>
  kable_styling(c('striped', 'hover'),
                latex_options = c('HOLD_position', 'scale_down')
                )
```

```{r}
#| eval: false
# Student views per enrollee by item (bar chart)
dat_temp <- dat |>
  group_by(Citation) |>
  summarize(Views.per.enrollee = sum(Total.views) / (sum(Course.enrollment) / n())
            )
ggplot(dat_temp,
       aes(x = Views.per.enrollee,
           y = Citation
           )
       ) +
  geom_col(fill = colors_penn$primary[1]) +
  labs(title = 'ALl views per student and course',
       x = 'View rate'
       ) +
  geom_text(aes(label = round(Views.per.enrollee, 2)),
            color = 'white',
            hjust = +1
            ) +
  theme_classic()
```

## Summary by term

```{r}
# Number of student views by term (table)
dat_temp <- dat |>
  group_by(Course.year, Course.term) |>
  summarize(Courses = length(unique(Course.code)),
            'Viewed\nitems' = length(unique(Citation)),
            'Minimum\nenrollment' = min(Course.enrollment),
            'Maximum\nenrollment' = max(Course.enrollment),
            Views = sum(Total.views)
            ) |>
  arrange(Course.year, Course.term)
names(dat_temp)[c(1,2)] <- c('Course\nyear', 'Course\nterm')
kbl(dat_temp,
    format = 'latex',
    caption = 'Number of student views by term',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE
    ) |>
  kable_styling(c('striped', 'hover'),
                latex_options = c('HOLD_position', 'scale_down')
                )
```

```{r}
#| eval: false
# Student views per citation by term (bar chart)
dat_temp <- dat |>
  group_by(Course.year, Course.term) |>
  summarize(Views.per.citation = sum(Total.views) / (sum(Course.enrollment) / n() / length(unique(Citation)))
            )
ggplot(dat_temp,
       aes(x = Views.per.citation,
           y = paste(Course.year, Course.term)
           )
       ) +
  geom_col(fill = colors_penn$primary[1]) +
  labs(title = 'Views per student and item',
       x = 'View rate',
       y = 'Term'
       ) +
  geom_text(aes(label = round(Views.per.citation, 2)),
            color = 'white',
            hjust = +1
            ) +
  theme_classic()
```

## Views by date

```{r}
# Hard-code presets for histogram
presets <- list(
  Term = c('Spring'),
  Year = c('2025'),
  Event = c('2025-04-05')
)
```

Include this section only as needed.

To do:

- Interpolate missing dates.
- Adjust vertical date label position so it's next to the line.

```{r}
# Views by date (histogram)
library(ggplot2)
dat_temp <- dat |>
  filter(Course.term %in% presets$Term & Course.year %in% presets$Year) |>
  group_by(Event.date) |>
  summarize(Views = sum(Total.views))
ggplot(dat_temp,
       aes(x = Event.date,
           y = Views
           )
       ) +
  geom_col(fill = colors_penn$primary[1]) +
  geom_vline(xintercept = presets$Event, linetype = "dashed", color = colors_penn$primary[2], size = .5) +
  annotate('text', x = presets$Event, y = 0, label = presets$Event, color = "black", hjust = -2.5, angle = 90) +
  scale_x_discrete(breaks = dat_temp$Event.date[seq(1, nrow(dat_temp), by = 14)]) +
  labs(title = 'Views by date',
       x = 'Date'
       ) +
  theme_classic()
```

