---
title: "Course Reserves Statistics"
author: "Kevin A Thomas"
format: html
server: shiny
---

# Demonstration template

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive documents allow readers to modify parameters and see the results immediately. Learn more about Shiny interactive documents at <https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change. This demonstrates how a standard R plot can be made interactive:

```{r}
sliderInput("bins", "Number of bins:", 
            min = 1, max = 50, value = 30)
plotOutput("distPlot")
```

```{r}
#| context: server
output$distPlot <- renderPlot({
   x <- faithful[, 2]  # Old Faithful Geyser data
   bins <- seq(min(x), max(x), length.out = input$bins + 1)
   hist(x, breaks = bins, col = 'darkgray', border = 'white',
        xlab = 'Waiting time to next eruption (in mins)',
        main = 'Histogram of waiting times')
})
```

# Course Reserves Statistics

```{r}
# Penn colors
colors_penn <- list(primary = c('011f5b', '990000'),
                    secondary = c('3479be', 'aa755', 'bbd86b', 'bd60a5'),
                    gray = c('555d6a', '757f8f', 'a7aeba', 'e1e4ec'))
```

```{r}
# Load data
dat <- read.csv('~/GitHub/course-reserves-stats/res/leganto-usage-views-2026-03-10.csv')
#Replace hard-coded filepath with a dynamic selector in a GUI
```

## Summary by item

```{r}
# Number of student views by item (table)
#Values: Sum of column values
#Rows: Citation
#Columns: Max of Course enrollment, Files downloaded, File views, Full text views, Total views
#Static filters: User role = 'Student'
#User-selectable filters: Course year, Course term, Course code
```

```{r}
# Percent of student views by item (bar chart)
#Values: (Sum of Total views) divided by (max of Course enrollment)
#Categories: Citation
#Static filters: User role = 'Student'
#User-selectable filters (shared with the table above): Course year, Course term, Course code
```

## Summary by term

```{r}
# Number of student views by term (table)
#Values: Sum of column values
#Rows: Combination of 'Course year' and 'Course term'
#Columns: Max of 'Course enrollment' per 'Course code', count of unique 'Course code', 'Files downloaded', 'File views', 'Full text views', 'Total views'
#Static filters: 'User role' = 'Student'
#User-selectable filters: 'Course code'
```

```{r}
# Percent of student views by term (bar chart)
#Values: (Sum of 'Total views') divided by (sum of (max of 'Course enrollment' per 'Course code'))
#Categories: Combination of 'Course year' and 'Course term'
#Static filters: 'User role' = 'Student'
#User-selectable filters (shared with the table above): 'Course code'
```
