---
title: "Course Reserves Statistics"
author: "Lippincott Library"
date: today
execute: 
  echo: false
  warning: false
  error: false
format: 
  pptx:
    reference-doc: "Lippincott Library Presentation Template 0525.potx"
---

```{r}
# Dependencies
library(dplyr)
library(kableExtra)
library(ggplot2)

# Penn colors
colors_penn <- list(primary = c('#011f5b', '#990000'),
                    secondary = c('#3479be', '#faa755', '#bbd86b', '#bd60a5'),
                    gray = c('#555d6a', '#757f8f', '#a7aeba', '#e1e4ec'))
```

```{r}
# Load and filter data (data file with relevant course codes only)
dat <-
  read.csv('~/GitHub/course-reserves-stats/res/leganto-usage-views-2026-05-18.csv') |>
  filter(User.role == 'Student')
dat$Citation <- as.character(dat$Citation)
dat$Reading.list <- as.character(dat$Reading.list)
```

```{r}
# Concatenate key course terms
term_min <- paste(dat$Course.term[dat$Event.date == min(dat$Event.date)][1],
                  dat$Course.year[dat$Event.date == min(dat$Event.date)][1])
term_max <- paste(dat$Course.term[dat$Event.date == max(dat$Event.date)][1],
                  dat$Course.year[dat$Event.date == max(dat$Event.date)][1])
```

## About the data

These data cover courses from `{r} if (term_max == term_min) term_min else paste(term_min, "through", term_max)`.

Items that were viewed zero times do not appear. Lists containing only items that were viewed zero times do not appear either.

## Item statistics: Overall

```{r}
dat_temp <- dat |>
  group_by(Reading.list) |>
  summarize(Citations = length(unique(Citation)),
            Views = sum(Total.views)
            ) |>
  arrange(Reading.list)
```

- Minimum items per list: `{r} min(dat_temp$Citations)`
- Mean items per list: `{r} round(mean(dat_temp$Citations), digits = 1)`
- Median items per list: `{r} median(dat_temp$Citations)`
- Maximum items per list: `{r} max(dat_temp$Citations)`

Number of lists: `{r} nrow(dat_temp)`


## Item statistics: By department

```{r}
dat_temp <- dat |>
  group_by(Reading.list, Processing.department) |>
  summarize(Citations = length(unique(Citation))
            ) |>
  group_by(Processing.department) |>
  summarize(Minimum = min(Citations),
            Mean = round(mean(Citations),
                         digits = 1),
            Median = median(Citations),
            Maximum = max(Citations)
            ) |>
  arrange(Processing.department)
kbl(dat_temp,
    format = 'markdown',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE,
    longtable = TRUE
    )
```

## Statistics by list

```{r}
##| tbl-cap: "Statistics by list"
dat_temp <- dat |>
  group_by(Reading.list) |>
  summarize(Courses = length(unique(Course.code)),
            'Median\nenrollment' = median(Course.enrollment),
            Citations = length(unique(Citation)),
            Views = sum(Total.views)
            ) |>
  arrange(Reading.list)
kbl(dat_temp,
#    format = 'latex',
    format = 'markdown',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE,
    longtable = TRUE
    )# |>
#  kable_styling(c('striped'),
#                latex_options = c('HOLD_position', 'repeat_header')
#                )
```

## ??

