---
title: "Course Reserves: List Statistics"
author: "Lippincott Library"
date: today
execute: 
  echo: false
  warning: false
  error: false
format: 
  pptx:
    reference-doc: "Lippincott Library Presentation Template 0525.potx"
---

```{r}
# Dependencies
library(dplyr)
library(kableExtra)
library(ggplot2)
library(readr)

# Penn colors
colors_penn <- list(primary = c('#011f5b', '#990000'),
                    secondary = c('#3479be', '#faa755', '#bbd86b', '#bd60a5'),
                    gray = c('#555d6a', '#757f8f', '#a7aeba', '#e1e4ec'))
```

```{r}
# Load, filter, and merge data
dat_usage <-
  read_csv('~/GitHub/course-reserves-stats/res/leganto-usage-views-2026-05-18.csv',
           name_repair = 'universal',
           col_types = cols(.default = 'c',
                            Event.date = 'D',
                            Course.enrollment = 'n',
                            Files.downloaded = 'n',
                            File.views = 'n',
                            Full.text.views = 'n',
                            Total.views = 'n'
                            )
           ) |>
  filter(User.role == 'Student')
dat_map <-
  read_csv('~/GitHub/course-reserves-stats/res/leganto-course-citations-2026-05-19.csv',
           name_repair = 'universal',
           col_types = cols(.default = 'c',
                            Modification.date = 'T'
                            )
           )
dat <- full_join(dat_usage, dat_map,
                 by = join_by(Course, Reading.list, Citation)
                 ) |>
  filter(Course %in% dat_usage$Course)
```

```{r}
# Concatenate key course terms
term_min <- paste(dat$Course.term[dat$Event.date == min(dat$Event.date)][1],
                  dat$Course.year[dat$Event.date == min(dat$Event.date)][1])
term_max <- paste(dat$Course.term[dat$Event.date == max(dat$Event.date)][1],
                  dat$Course.year[dat$Event.date == max(dat$Event.date)][1])
```

## About the data

These data cover courses from `{r} if (term_max == term_min) term_min else paste(term_min, "through", term_max)`.

Citations that were viewed zero times do not appear and do not roll up to larger groupings, like lists.

```{r}
dat_temp <- dat |>
  group_by(Reading.list) |>
  summarize(Citations = length(unique(Citation)),
            Views = sum(Total.views)
            ) |>
  arrange(Reading.list)
```

## List size across all `{r} nrow(dat_temp)` lists

- Smallest list: `{r} min(dat_temp$Citations)`
- Median citations per list: `{r} median(dat_temp$Citations)`
- Mean citations per list: `{r} round(mean(dat_temp$Citations), digits = 1)`
- Largest list: `{r} max(dat_temp$Citations)`

## List size by department

```{r}
dat_temp <- dat |>
  group_by(Processing.department, Reading.list) |>
  summarize(Citations = length(unique(Citation))
            ) |>
  group_by(Processing.department) |>
  summarize(Lists = n(),
            'Smallest list' = min(Citations),
            'Median citations' = median(Citations),
            'Mean citations' = round(mean(Citations),
                         digits = 1),
            'Largest list' = max(Citations)
            ) |>
  arrange(Processing.department)
kbl(dat_temp,
    format = 'markdown',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE,
    longtable = TRUE
    )
```

## List size by department

Box plots display the median list size (center line), middle half of the lists (boxed area), expected variation in list size (whiskers), and unexpected list sizes (dots).

```{r}
dat_temp <- dat |>
  group_by(Processing.department, Reading.list) |>
  summarize(Citations = length(unique(Citation))
            ) |>
  group_by(Processing.department)
for(dept in unique(dat_temp$Processing.department)){
  dat_temp$Processing.department[dat_temp$Processing.department == dept] <- paste0(dept, " (", sum(dat_temp$Processing.department == dept), ")")
}
ggplot(dat_temp,
       aes(x = Citations, y = Processing.department)
    ) +
  geom_boxplot() +
  labs(y = "Processing Department (N)")# +
#  geom_jitter()
```

## Large lists (dots from the box plot)

```{r}
dat_temp <- dat |>
  group_by(Processing.department, Reading.list) |>
  summarize(Course = Course.name[1],
            Citations = length(unique(Citation))
            ) |>
  arrange(desc(Citations),
          Processing.department,
          Course,
          Reading.list
          )
dat_temp$Include <- FALSE
# Define large lists as outliers by Tukey's method: Q3 + 1.5*IQR
for(dept in unique(dat_temp$Processing.department)){
  data_dept <- dat_temp$Citations[dat_temp$Processing.department == dept]
  Q1 <- quantile(data_dept, .25)
  Q3 <- quantile(data_dept, .75)
  limit <- Q3 + ((Q3 - Q1) * 1.5)
  dat_temp$Include[dat_temp$Processing.department == dept & dat_temp$Citations > limit] <- TRUE
}
dat_temp <- subset(dat_temp, Include, select = -Include)
kbl(dat_temp,
    format = 'markdown',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE,
    longtable = TRUE
    )
```

## Citations in multiple lists

```{r}
dat_temp <- dat |>
  group_by(Citation) |>
  summarize(Lists = length(unique(Reading.list))) |>
  subset(Lists > 1) |>
  arrange(Lists, desc(Citation))
```

`{r} if(nrow(dat_temp) > 0){}else{'No repeated citations'}`

```{r}
kbl(dat_temp,
    format = 'markdown',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE,
    longtable = TRUE
    )
```

## Statistics by list - Do not keep as-is

```{r}
##| tbl-cap: "Statistics by list"
dat_temp <- dat |>
  group_by(Reading.list) |>
  summarize(Courses = length(unique(Course.code)),
            Course = Course.name[1],
            'Course enrollment' = median(Course.enrollment),
            Citations = length(unique(Citation)),
            Views = sum(Total.views)
            ) |>
  arrange(Reading.list)
kbl(dat_temp,
#    format = 'latex',
    format = 'markdown',
    col.names = linebreak(names(dat_temp)),
    escape = FALSE,
    longtable = TRUE
    )# |>
#  kable_styling(c('striped'),
#                latex_options = c('HOLD_position', 'repeat_header')
#                )
```
